### 2016-05-04 p01-ch01-2012/paper.tex
Fix invalid Figure references
Fix ref/label for Chapter 01

\section{Introduction} \label{sec:introduction}
One of the major challenges in the design of programming
languages is to provide high-level support for  multi-core and cloud  applications which are becoming increasingly important.
Both multi-core and cloud  applications require an explicit and precise treatment of non-functional properties, e.g., resource requirements.
On the cloud, services execute in the
context of virtual resources, and the amount of resources actually available to a service is subject to change.
Multi-core applications require techniques 
to help the programmer optimally use potentially many cores. At
the operating system level, resource management is greatly affected by scheduling which
is largely beyond the control of most existing high-level programming languages.
Therefore, for optimal use of the available resources, we cannot avoid leveraging
scheduling and performance related issues from the underlying operating system
to the application level. However, the very nature of high-level languages is to
provide suitable abstractions that hide
implementation details from the programmer. The main challenge in designing programming languages
for  multi-core and cloud  applications
is to find a balance between these two conflicting requirements.

@@

Section \ref{sec:creol} first provides an overview of the Creol language with application-level scheduling. 
In Section \ref{sec:compiler}, we elaborate on the design  of the tool-set and the
prototype. The use of the tool-set is exemplified by a case study in Section \ref{sec:caseStudy}.
Section \ref{sec:relwork} summarizes the related work. 
Finally, we conclude in Section \ref{sec:conclusion}.

%\section{Preliminaries: Creol} \label{sec:creol}
%\input{Creol}

@@

\section{Application-Level Scheduling}
\label{sec:creol}

@@

\begin{lstlisting}[float=t, label=lst:exres, caption=Exclusive Resource in Creol %, multicols=2
]
interface Resource begin
	(* \label{lst:priority_range} *)(*
	%with User  \label{lst:with} 
	*)
		op request()
		op release()
end

class ExclusiveResource implements Resource begin
	var taken := false;
	(* \label{lst:priority_class} *)
	op request () == 
		await (*$\sim$*)taken; 			(* \label{lst:await_taken} *)
		taken := true;
	op release () ==
		taken := false
end
\end{lstlisting}

On the server side, an interface may introduce a \textit{priority
range} that is available to all clients. For instance, in Line
\ref{lst:priority_range} of {\footnotesize\texttt{Resource}} in Listing
\ref{lst:exres}, we can define a priority range:

@@

Scheduling \textit{only} on the basis of client-side priority requirements is too restrictive.
For example, if there are many \ftt{request} messages with high priorities and a low priority \ftt{release}, the server might block as it would fail to schedule the \ftt{release}.
In this particular example, we can solve this problem by allowing the servers to
prioritize their methods.
This involves a declaration of a priority range generally depending on the structure of the class. 
In our example, assuming a range 0..2 added in Line \ref{lst:priority_class}, this requires changing the method signatures in the {\ftt{Exclusive\-Resource}} class:

@@

This gives \ftt{release}  a higher priority over \ftt{request}, because by default, the smaller values indicate higher priorities.
Furthermore, the server may also introduce priorities on certain 
characteristics of a method invocation such as the kind of ``release statement''
being executed. 
For example, a process waiting at Line \ref{lst:await_taken} could be given a higher priority over new requests by assigning a smaller value:

@@

% 
% This is an example of {\em method level} priority specification, which is specified inside a class together with the definition of a method. 
% Method definitions without an explicit priority are given a default priority value. 
% %As discussed in Section \ref{sec:conclusion}, 
% Other possible levels of priority include method invocation level, specified by the caller, or interface level priorities. 
% %

@@

Note that the abstract priority manager in general does \textit{not} completely
fix the ``choice'' of which method invocation to execute. 
In our tool-set, we include an extensible library of predefined scheduling policies such as strong or weak fairness that further refine the application-specific multi-level priority scheduling. 
The policies provided by the library are available to the user to annotate the classes.
We may declare a scheduling policy for the \ftt{ExclusiveResource} class by adding at Line
\ref{lst:priority_class} in Listing \ref{lst:exres}: 
%
\begin{quote}
\lstinline{scheduling policy StronglyFairScheduler;} 
\end{quote}

@@

\section{Tool Architecture} \label{sec:compiler}

\begin{figure}[t]
\begin{center}
  %\vspace{-.1cm}
  \includegraphics[scale=0.36]{figs/manticore-concurrency-v4}
  \caption{\Crisp Architecture: Structural Overview}
  \label{fig:api}
\end{center}
\end{figure}

We have implemented a tool to translate Creol programs into Java programs for execution, called {\em \Crisp} (Creolized Interacting Scheduling Policies).
\Crisp provides a one-to-one mapping from Creol classes and methods to their equivalent Java constructs. 
In order to implement active objects in Creol, we use the \jtt{java.util.concurrent} package (see Figure \ref{fig:api}).
Each active object consists of an instance of a process store  and an execution manager to store and execute the method invocations.

@@

\subsection{A New Method Invocation}
\label{sec:add}

\begin{figure}[t]
%\begin{center}
\includegraphics[scale=0.36]{figs/seq-diag-1}
~\\[-3.5cm]ADD:\\[2.5cm]
\caption[New MethodInvocation]{Adding the new MethodInvocations are performed on the Client side.}
\label{fig:add}
%\end{center}
\end{figure}

@@

In option $1$, the server may be busy doing something else. Therefore,
in this case the client must wait until the server is free,  which is
against the asynchronous nature of communication. In Option $2$, the
Java implementation of each active object exposes its process store as
an immutable public property. Thus, the actual code for adding the new
method invocation instance is run in the execution thread of the
client. 
We adopt the second approach as depicted in Figure \ref{fig:add}.
At any time, there can be concurrent clients that are
storing instances of \ftt{MethodInvocation} into the server's process store, but since
the process store implementation encapsulates the mechanisms for concurrency and
data safety, the clients have no concern on data synchronization and concurrency
issues such as mutual exclusion.
The method or policy used to store the method invocation in the process store of
the server is totally up to the server's process store implementation details.
%Storing a method invocation in the process store of the server takes place in
%the client side thread of execution. Listing \ref{lst:jfib_method_fibonacci}
%presents a sample generated Java code for an asynchronous method call in Listing
%\ref{lst:cfib}.

@@

On the server side of the story,  an active object 
repeatedly fetches an instance of method invocation from its process
store for execution (cf. Fig \ref{fig:take}).
The process store uses its instance of
\ftt{SchedulingManager} to choose one of the method invocations. 
\Crisp has some predefined scheduling policies that can be used as scheduling managers;
nevertheless, new scheduling policies can be easily developed
and customized based on the requirements by the user. 
%Listing \ref{lst:aoloop} presents the main loop for the active object.

\begin{figure}[t]
~\\TAKE:\\
\begin{center}
%\vspace{-1.3cm}
  \includegraphics[scale=0.3]{figs/seq-diag-2}
% \vspace{-.7cm}
\caption[Policy-based selection of a MethodInvocation]{An active object selects a method invocation based on its local scheduling policy. After a method finishes execution, the whole scenario is repeated.}
\end{center}
\label{fig:take}
\end{figure}

@@

Every method invocation may carry different levels of priority information,
e.g., a server side priority assigned to the method or a
client side priority. 
The \ftt{PriorityManager} 
%is an interface the implementations of which are supposed to 
provides a function to determine and
resolve a final priority value in case there are different levels of priorities
specified for a method invocation. 
% This feature is used, if specified, by the
% process store to compute the ``resolved priority'' of the method invocation
% based on different levels of priorities specified and then add it to the
% storage. 
%An example of a user-defined priority manager is given in Section \ref{sec:caseStudy}.
Postponing the act of resolving priorities to this point rather than when inserting new processes to the store enables us to handle dynamic priorities.

@@

As shown in Figure \ref{fig:execute}, the thread factory creates threads of type  {\ftt{InterruptibleProcess}}.
The execution manager thread blocks as soon as it starts the interruptible process which executes the associated  method invocation. 
If the method releases the processor before completion, it will be added back to the process store as explained in Section \ref{sec:add}.
When a suspended method invocation is resumed, the execution manager skips the creation of a new thread and  reuses the one that was assigned to the method invocation before.

\begin{figure}[t]
~\\EXECUTE:\\
\begin{center}
% \vspace{-1.3cm}
 \includegraphics[scale=0.35]{figs/seq-diag-3}
% \vspace{-.7cm}
\caption[Execution of a MethodInvocation]{A method invocation is executed in an interruptible process. The execution manager thread is blocked while the interruptible process is running.}
\end{center}
\label{fig:execute}
\end{figure}

@@

\section{Case Study} \label{sec:caseStudy}

@@

We parallelize this algorithm by creating active objects that run in parallel.
The numbers are passed asynchronously as a parameter to the \ftt{divide} message. 
Correctness of the parallel algorithm essentially depends on the numbers being processed in increasing order.
For example, if object \ftt{two} processes 9 before 3, it will erroneously treat 9 as a prime, because 3 is not there yet to eliminate 9. 
%Such a scenario is possible if the scheduling of the tasks in the \ftt{Prime} objects is not specified. 
%
To avoid erroneous behavior, we use the actual parameter $n$ in \ftt{divide} method to define its priority level, too (see line \ref{line:priority}). 
As a result, every invocation of this method generates a process with a priority equal to its parameter.
The default scheduling policy for objects always selects a process for execution that has the smallest priority value. 
This guarantees that the numbers sent to a \ftt{Prime} object are processed exactly in increasing order.

@@

class Prime(p: Int) implements IPrime begin (*\label{lst:priority}*)
	var next: IPrime
	op divide(n: Int) priority (n) == (*\label{line:priority}*)
		if (n % p) /= 0 then 
			if next /= null then
				!next.divide(n)
			else
				next := new Prime(n)
		end end
end
\end{lstlisting}
% 
% class Test
% begin
% 	var g : Sieve
% 	op run == g := new Sieve
% end

@@

The first interesting observation was that \Crisp prime sieve utilizes all the CPU's on the underlying hardware as much as possible during  execution. 
This can be seen in Figure \ref{fig:multi-cpu-1} which shows the CPU usage. 
Both CPUs are fairly in use while running this program.
Figure \ref{fig:prime_sieve_cacoj} depicts the results of the monitoring of the parallel prime sieve in \Crisp using Visual VM tool. 
It depicts the number of threads generated for the program.
This shows that \Crisp can handle a massive number of concurrent tasks.

One interesting feature of \Crisp is that the execution of any program under \Crisp can constantly utilize the \emph{minimum} memory that can be allocated for each thread in JVM (thread stack). 
In JVM, the size of thread stack can be configured using \ftt{-Xss} option for every run. 
To demonstrate this feature of \Crisp, we collected the minimum stack size needed for every program run in Table \ref{tbl:cmp-thread-stack}. 
All \Crisp runs use the minimum thread stack size of \ftt{64k} that is possible for the JVM.
On the contrary, the stack size required for the sequential version of the sieve program increases by the number of primes detected. 
This is also expected because of the long chain of method calls in the sequential sieve.

% The results depict that the sequential version, through time, increasingly required memory allocation for the thread stack, however, \Crisp performs on a constant size of thread stack. 
% Table \ref{tbl:cmp-thread-stack} summarizes the thread stack allocation on different runs.

\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c c c c}
$max$ & $10000$ & $20000$ & $30000$ & $50000$ & $100000$ \\ \thickhline
Sequential & 64k & 72k & 96k & 160k & 190k \\ 
\Crisp & 64k & 64k & 64k & 64k & 64k \\ \thickhline
\end{tabular}
\caption{Thread stack allocated for different executions}
\label{tbl:cmp-thread-stack}
\end{center}
\end{table}

Having the constant thread stack size feature, \Crisp provides another interesting feature. It can handle huge number of thread generation if required. Table \ref{tbl:cmp-thread-num-gen} summarizes the thread generation data for parallel prime sieve in \Crisp. It shows scalability of \Crisp as $p$ rises for parallel prime sieve.
\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c}
$max$ & Live Peak & Total \\ \thickhline
$10000$ & 817 & 540591 \\
$20000$ & 1468 & 1854067 \\
$30000$ & 2204 & 4054814 \\
$50000$ & 3707 & 11852985 \\ \thickhline
\end{tabular}
\caption{Number of live threads and total threads created for different runs of parallel prime sieve}
\label{tbl:cmp-thread-num-gen}
\end{center}
\end{table}

As the results show the use of Java threads is costly; first, \Crisp
does not need much of the memory allocated to each thread and, second,
the context switch cost is higher for larger memory allocation. In line
with this, JVM uses a \emph{one-to-one} mapping from an
application-level Java thread to a native OS-level thread. 
In the current setting, the context switch of the threads are in
the OS level. When the context switch is taken to the application
level, we leverage the performance issue from the OS level to the
application level. We further discuss this in Section
\ref{sec:conclusion}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.5]{figs/p50000vvm}
\caption{Increasing parallelism in \Crisp for Prime Sieve}
\label{fig:prime_sieve_cacoj}
% \end{center}
% \end{figure}
% 
% \begin{figure}[t]
% \begin{center}
  \includegraphics[scale=0.44]{figs/p50000sysmon} 
  \caption{Utilizing both CPUs with Prime Sieve in \Crisp}
  \label{fig:multi-cpu-1}
\end{center}
\end{figure}

\section{Related Work} \label{sec:relwork}
\input{related2}

\section{Conclusion} \label{sec:conclusion}
In this paper, we proposed \Crisp as an implementation scheme for
application-level scheduling of active objects. 
\Crisp first introduces asynchronous message passing with fine-grain
priority management and scheduling of messages. 
Additionally, it introduces a Creol to Java compiler that translates
the active objects in Creol into an equivalent Java application. 
\Crisp compiler seamlessly integrates Java class libraries into Creol
including data types that turns Creol from a modeling language to a
fully-fledged one in the hands of the programmer. 
% \Crisp compiler takes advantage of template engines; i.e. the code
% generation templates can easily be customized in the direction of the
% requirements of the problems at hand.

The \jtt{java.util.\-con\-current} package provides useful API for
concurrent programming. Java futures facilitate modeling
asynchronous message passing. However, for processor release
points, we had to preserve threads (using \jtt{Interruptible\-Process})
to allow continuations which leads to their OS-level context switching
that is costly. Moreover, we were tightly directed to use the
out-of-the-box \jtt{Executor\-Service} which is limitedly extensible. We
had no control over the scheduling mechanisms of the internal queue used
in the service implementations. Thus, we needed to re-implement some of
the concepts. Through prototyping \Crisp, we learned that
there are two major
challenges ahead. Firstly, we need to integrate continuations into Java
using a many-messages-to-one-thread mapping model. Secondly, we need
complete control over scheduling of messages and threads in
\jtt{Executor\-Service}'s internal queue. Table \ref{tbl:eval}
summarizes this discussion.

\begin{table}[h]\small
\begin{center}
\begin{tabular}{p{1.5cm} p{2cm} p{2cm} p{1.5cm}}
 & Asynchronous Communication & Processor Release Point & Scheduling \\
\thickhline
Modeling & \ding{51} & \ding{51} & \ding{53} \\
Performance & \ding{51} & \ding{53} & \ding{51} \\ \thickhline
 & Java Futures & Interruptible Process & Executor Service \\
\end{tabular}
\caption{Overview of evaluation of challenges}
\label{tbl:eval}
\end{center}
\end{table}

>>>>

\section{Introduction} \label{ch01:sec:introduction}
One of the major challenges in the design of programming
languages is to provide high-level support for  multi-core and cloud  applications which are becoming increasingly important.
Both multi-core and cloud  applications require an explicit and precise treatment of non-functional properties, e.g., resource requirements.
On the cloud, services execute in the
context of virtual resources, and the amount of resources actually available to a service is subject to change.
Multi-core applications require techniques 
to help the programmer optimally use potentially many cores. At
the operating system level, resource management is greatly affected by scheduling which
is largely beyond the control of most existing high-level programming languages.
Therefore, for optimal use of the available resources, we cannot avoid leveraging
scheduling and performance related issues from the underlying operating system
to the application level. However, the very nature of high-level languages is to
provide suitable abstractions that hide
implementation details from the programmer. The main challenge in designing programming languages
for  multi-core and cloud  applications
is to find a balance between these two conflicting requirements.

@@

Section \ref{ch01:sec:creol} first provides an overview of the Creol language with application-level scheduling. 
In Section \ref{ch01:sec:compiler}, we elaborate on the design  of the tool-set and the
prototype. The use of the tool-set is exemplified by a case study in Section \ref{ch01:sec:caseStudy}.
Section \ref{ch01:sec:relwork} summarizes the related work. 
Finally, we conclude in Section \ref{ch01:sec:conclusion}.

%\section{Preliminaries: Creol} \label{ch01:sec:creol}
%\input{Creol}

@@

\section{Application-Level Scheduling}
\label{ch01:sec:creol}

@@

\begin{lstlisting}[float=t, label=lst:exres, caption=Exclusive Resource in Creol %, multicols=2
]
interface Resource begin
	(* \label{ch01:lst:priority_range} *)(*
	%with User  \label{ch01:lst:with} 
	*)
		op request()
		op release()
end

class ExclusiveResource implements Resource begin
	var taken := false;
	(* \label{ch01:lst:priority_class} *)
	op request () == 
		await (*$\sim$*)taken; 			(* \label{ch01:lst:await_taken} *)
		taken := true;
	op release () ==
		taken := false
end
\end{lstlisting}

On the server side, an interface may introduce a \textit{priority
range} that is available to all clients. For instance, in Line
\ref{ch01:lst:priority_range} of {\footnotesize\texttt{Resource}} in Listing
\ref{ch01:lst:exres}, we can define a priority range:

@@

Scheduling \textit{only} on the basis of client-side priority requirements is too restrictive.
For example, if there are many \ftt{request} messages with high priorities and a low priority \ftt{release}, the server might block as it would fail to schedule the \ftt{release}.
In this particular example, we can solve this problem by allowing the servers to
prioritize their methods.
This involves a declaration of a priority range generally depending on the structure of the class. 
In our example, assuming a range 0..2 added in Line \ref{ch01:lst:priority_class}, this requires changing the method signatures in the {\ftt{Exclusive\-Resource}} class:

@@

This gives \ftt{release}  a higher priority over \ftt{request}, because by default, the smaller values indicate higher priorities.
Furthermore, the server may also introduce priorities on certain 
characteristics of a method invocation such as the kind of ``release statement''
being executed. 
For example, a process waiting at Line \ref{ch01:lst:await_taken} could be given a higher priority over new requests by assigning a smaller value:

@@

% 
% This is an example of {\em method level} priority specification, which is specified inside a class together with the definition of a method. 
% Method definitions without an explicit priority are given a default priority value. 
% %As discussed in Section \ref{ch01:sec:conclusion}, 
% Other possible levels of priority include method invocation level, specified by the caller, or interface level priorities. 
% %

@@

Note that the abstract priority manager in general does \textit{not} completely
fix the ``choice'' of which method invocation to execute. 
In our tool-set, we include an extensible library of predefined scheduling policies such as strong or weak fairness that further refine the application-specific multi-level priority scheduling. 
The policies provided by the library are available to the user to annotate the classes.
We may declare a scheduling policy for the \ftt{ExclusiveResource} class by adding at Line
\ref{ch01:lst:priority_class} in Listing \ref{ch01:lst:exres}: 
%
\begin{quote}
\lstinline{scheduling policy StronglyFairScheduler;} 
\end{quote}

@@

\section{Tool Architecture} \label{ch01:sec:compiler}

\begin{figure}[t]
\begin{center}
  %\vspace{-.1cm}
  \includegraphics[scale=0.36]{figs/manticore-concurrency-v4}
  \caption{\Crisp Architecture: Structural Overview}
  \label{ch01:fig:api}
\end{center}
\end{figure}

We have implemented a tool to translate Creol programs into Java programs for execution, called {\em \Crisp} (Creolized Interacting Scheduling Policies).
\Crisp provides a one-to-one mapping from Creol classes and methods to their equivalent Java constructs. 
In order to implement active objects in Creol, we use the \jtt{java.util.concurrent} package (see Figure \ref{ch01:fig:api}).
Each active object consists of an instance of a process store  and an execution manager to store and execute the method invocations.

@@

\subsection{A New Method Invocation}
\label{ch01:sec:add}

\begin{figure}[t]
%\begin{center}
\includegraphics[scale=0.36]{figs/seq-diag-1}
~\\[-3.5cm]ADD:\\[2.5cm]
\caption[New MethodInvocation]{Adding the new MethodInvocations are performed on the Client side.}
\label{ch01:fig:add}
%\end{center}
\end{figure}

@@

In option $1$, the server may be busy doing something else. Therefore,
in this case the client must wait until the server is free,  which is
against the asynchronous nature of communication. In Option $2$, the
Java implementation of each active object exposes its process store as
an immutable public property. Thus, the actual code for adding the new
method invocation instance is run in the execution thread of the
client. 
We adopt the second approach as depicted in Figure \ref{ch01:fig:add}.
At any time, there can be concurrent clients that are
storing instances of \ftt{MethodInvocation} into the server's process store, but since
the process store implementation encapsulates the mechanisms for concurrency and
data safety, the clients have no concern on data synchronization and concurrency
issues such as mutual exclusion.
The method or policy used to store the method invocation in the process store of
the server is totally up to the server's process store implementation details.
%Storing a method invocation in the process store of the server takes place in
%the client side thread of execution. Listing \ref{ch01:lst:jfib_method_fibonacci}
%presents a sample generated Java code for an asynchronous method call in Listing
%\ref{ch01:lst:cfib}.

@@

On the server side of the story,  an active object 
repeatedly fetches an instance of method invocation from its process
store for execution (cf. Fig \ref{ch01:fig:take:next}).
The process store uses its instance of
\ftt{SchedulingManager} to choose one of the method invocations. 
\Crisp has some predefined scheduling policies that can be used as scheduling managers;
nevertheless, new scheduling policies can be easily developed
and customized based on the requirements by the user. 
%Listing \ref{ch01:lst:aoloop} presents the main loop for the active object.

\begin{figure}[t]
~\\TAKE:\\
\begin{center}
%\vspace{-1.3cm}
  \includegraphics[scale=0.3]{figs/seq-diag-2}
% \vspace{-.7cm}
\end{center}
\caption[Policy-based selection of a MethodInvocation]{An active object selects a method invocation based on its local scheduling policy. After a method finishes execution, the whole scenario is repeated.}
\label{ch01:fig:take:next}
\end{figure}

@@

Every method invocation may carry different levels of priority information,
e.g., a server side priority assigned to the method or a
client side priority. 
The \ftt{PriorityManager} 
%is an interface the implementations of which are supposed to 
provides a function to determine and
resolve a final priority value in case there are different levels of priorities
specified for a method invocation. 
% This feature is used, if specified, by the
% process store to compute the ``resolved priority'' of the method invocation
% based on different levels of priorities specified and then add it to the
% storage. 
%An example of a user-defined priority manager is given in Section \ref{ch01:sec:caseStudy}.
Postponing the act of resolving priorities to this point rather than when inserting new processes to the store enables us to handle dynamic priorities.

@@

As shown in Figure \ref{ch01:fig:execute}, the thread factory creates threads of type  {\ftt{InterruptibleProcess}}.
The execution manager thread blocks as soon as it starts the interruptible process which executes the associated  method invocation. 
If the method releases the processor before completion, it will be added back to the process store as explained in Section \ref{ch01:sec:add}.
When a suspended method invocation is resumed, the execution manager skips the creation of a new thread and  reuses the one that was assigned to the method invocation before.

\begin{figure}[t]
~\\EXECUTE:\\
\begin{center}
% \vspace{-1.3cm}
 \includegraphics[scale=0.35]{figs/seq-diag-3}
% \vspace{-.7cm}
\end{center}
\caption[Execution of a MethodInvocation]{A method invocation is executed in an interruptible process. The execution manager thread is blocked while the interruptible process is running.}
\label{ch01:fig:execute}
\end{figure}

@@

\section{Case Study} \label{ch01:sec:caseStudy}

@@

We parallelize this algorithm by creating active objects that run in parallel.
The numbers are passed asynchronously as a parameter to the \ftt{divide} message. 
Correctness of the parallel algorithm essentially depends on the numbers being processed in increasing order.
For example, if object \ftt{two} processes 9 before 3, it will erroneously treat 9 as a prime, because 3 is not there yet to eliminate 9. 
%Such a scenario is possible if the scheduling of the tasks in the \ftt{Prime} objects is not specified. 
%
To avoid erroneous behavior, we use the actual parameter $n$ in \ftt{divide} method to define its priority level, too (see line \ref{ch01:line:priority}). 
As a result, every invocation of this method generates a process with a priority equal to its parameter.
The default scheduling policy for objects always selects a process for execution that has the smallest priority value. 
This guarantees that the numbers sent to a \ftt{Prime} object are processed exactly in increasing order.

@@

class Prime(p: Int) implements IPrime begin (*\label{ch01:lst:priority}*)
	var next: IPrime
	op divide(n: Int) priority (n) == (*\label{ch01:line:priority}*)
		if (n % p) /= 0 then 
			if next /= null then
				!next.divide(n)
			else
				next := new Prime(n)
		end end
end
\end{lstlisting}
% 
% class Test
% begin
% 	var g : Sieve
% 	op run == g := new Sieve
% end

@@

The first interesting observation was that \Crisp prime sieve utilizes all the CPU's on the underlying hardware as much as possible during  execution. 
This can be seen in Figure \ref{ch01:fig:multi-cpu-1} which shows the CPU usage. 
Both CPUs are fairly in use while running this program.
Figure \ref{ch01:fig:prime_sieve_cacoj} depicts the results of the monitoring of the parallel prime sieve in \Crisp using Visual VM tool. 
It depicts the number of threads generated for the program.
This shows that \Crisp can handle a massive number of concurrent tasks.

One interesting feature of \Crisp is that the execution of any program under \Crisp can constantly utilize the \emph{minimum} memory that can be allocated for each thread in JVM (thread stack). 
In JVM, the size of thread stack can be configured using \ftt{-Xss} option for every run. 
To demonstrate this feature of \Crisp, we collected the minimum stack size needed for every program run in Table \ref{ch01:tbl:cmp-thread-stack}. 
All \Crisp runs use the minimum thread stack size of \ftt{64k} that is possible for the JVM.
On the contrary, the stack size required for the sequential version of the sieve program increases by the number of primes detected. 
This is also expected because of the long chain of method calls in the sequential sieve.

% The results depict that the sequential version, through time, increasingly required memory allocation for the thread stack, however, \Crisp performs on a constant size of thread stack. 
% Table \ref{ch01:tbl:cmp-thread-stack} summarizes the thread stack allocation on different runs.

\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c c c c}
$max$ & $10000$ & $20000$ & $30000$ & $50000$ & $100000$ \\ \thickhline
Sequential & 64k & 72k & 96k & 160k & 190k \\ 
\Crisp & 64k & 64k & 64k & 64k & 64k \\ \thickhline
\end{tabular}
\caption{Thread stack allocated for different executions}
\label{ch01:tbl:cmp-thread-stack}
\end{center}
\end{table}

Having the constant thread stack size feature, \Crisp provides another interesting feature. It can handle huge number of thread generation if required. Table \ref{ch01:tbl:cmp-thread-num-gen} summarizes the thread generation data for parallel prime sieve in \Crisp. It shows scalability of \Crisp as $p$ rises for parallel prime sieve.
\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c}
$max$ & Live Peak & Total \\ \thickhline
$10000$ & 817 & 540591 \\
$20000$ & 1468 & 1854067 \\
$30000$ & 2204 & 4054814 \\
$50000$ & 3707 & 11852985 \\ \thickhline
\end{tabular}
\caption{Number of live threads and total threads created for different runs of parallel prime sieve}
\label{ch01:tbl:cmp-thread-num-gen}
\end{center}
\end{table}

As the results show the use of Java threads is costly; first, \Crisp
does not need much of the memory allocated to each thread and, second,
the context switch cost is higher for larger memory allocation. In line
with this, JVM uses a \emph{one-to-one} mapping from an
application-level Java thread to a native OS-level thread. 
In the current setting, the context switch of the threads are in
the OS level. When the context switch is taken to the application
level, we leverage the performance issue from the OS level to the
application level. We further discuss this in Section
\ref{ch01:sec:conclusion}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.5]{figs/p50000vvm}
\caption{Increasing parallelism in \Crisp for Prime Sieve}
\label{ch01:fig:prime_sieve_cacoj}
% \end{center}
% \end{figure}
% 
% \begin{figure}[t]
% \begin{center}
  \includegraphics[scale=0.44]{figs/p50000sysmon} 
  \caption{Utilizing both CPUs with Prime Sieve in \Crisp}
  \label{ch01:fig:multi-cpu-1}
\end{center}
\end{figure}

\section{Related Work} \label{ch01:sec:relwork}
\input{related2}

\section{Conclusion} \label{ch01:sec:conclusion}
In this paper, we proposed \Crisp as an implementation scheme for
application-level scheduling of active objects. 
\Crisp first introduces asynchronous message passing with fine-grain
priority management and scheduling of messages. 
Additionally, it introduces a Creol to Java compiler that translates
the active objects in Creol into an equivalent Java application. 
\Crisp compiler seamlessly integrates Java class libraries into Creol
including data types that turns Creol from a modeling language to a
fully-fledged one in the hands of the programmer. 
% \Crisp compiler takes advantage of template engines; i.e. the code
% generation templates can easily be customized in the direction of the
% requirements of the problems at hand.

The \jtt{java.util.\-con\-current} package provides useful API for
concurrent programming. Java futures facilitate modeling
asynchronous message passing. However, for processor release
points, we had to preserve threads (using \jtt{Interruptible\-Process})
to allow continuations which leads to their OS-level context switching
that is costly. Moreover, we were tightly directed to use the
out-of-the-box \jtt{Executor\-Service} which is limitedly extensible. We
had no control over the scheduling mechanisms of the internal queue used
in the service implementations. Thus, we needed to re-implement some of
the concepts. Through prototyping \Crisp, we learned that
there are two major
challenges ahead. Firstly, we need to integrate continuations into Java
using a many-messages-to-one-thread mapping model. Secondly, we need
complete control over scheduling of messages and threads in
\jtt{Executor\-Service}'s internal queue. Table \ref{ch01:tbl:eval}
summarizes this discussion.

\begin{table}[h]\small
\begin{center}
\begin{tabular}{p{1.5cm} p{2cm} p{2cm} p{1.5cm}}
 & Asynchronous Communication & Processor Release Point & Scheduling \\
\thickhline
Modeling & \ding{51} & \ding{51} & \ding{53} \\
Performance & \ding{51} & \ding{53} & \ding{51} \\ \thickhline
 & Java Futures & Interruptible Process & Executor Service \\
\end{tabular}
\caption{Overview of evaluation of challenges}
\label{ch01:tbl:eval}
\end{center}
\end{table}
